Algebraic Groups\vspace{-1ex}

### 2019-10-26 Algebraic Geometry/Algebraic-Geometry.tex
update

\subsection{Important Morphisms}
\begin{defn}
	A morphism $f:X\to Y$ of varieties is 
	\begin{itemize}
		\item \textbf{affine} if there exists an affine covering $U_i\subseteq Y$ such that $f^{-1}(U_i)\subseteq X$ is affine.
		\item \textbf{finite} if there exists an affine covering $U_i\subseteq Y$ such that $f^{-1}(U_i)$ is affine and 
		\[\O_Y(U_i)\to\O_X(f^{-1}(U_i))\]
		is a finite ring extension.
	\end{itemize}
\end{defn}
\begin{prop}
	$f:X\to Y$ is affine (resp. finite) if and only if for all $U\subseteq Y$ affine, $f^{-1}(U)$ is affine (resp. $f^{-1}(U)$ is affine 
	and $\O_X(U)\to \O_X(f^{-1}(U))$ is finite).
\end{prop}

>>>>

\subsection{Important Morphisms}
\begin{defn}
	A morphism $f:X\to Y$ of varieties is 
	\begin{itemize}
		\item \textbf{affine} if there exists an affine covering $U_i\subseteq Y$ such that $f^{-1}(U_i)\subseteq X$ is affine.
		\item \textbf{finite} if there exists an affine covering $U_i\subseteq Y$ such that $f^{-1}(U_i)$ is affine and 
		\[\O_Y(U_i)\to\O_X(f^{-1}(U_i))\]
		is a finite ring extension (finitely generated as a module over the base).
	\end{itemize}
\end{defn}
\begin{prop}
	$f:X\to Y$ is affine (resp. finite) if and only if for all $U\subseteq Y$ affine, $f^{-1}(U)$ is affine (resp. $f^{-1}(U)$ is affine 
	and $\O_X(U)\to \O_X(f^{-1}(U))$ is finite).
\end{prop}

\section{October 25th, 2019}
Another takeaway from this class: \textit{algebraic geometry is relative and not absolute.} One way we are starting to dip our toes into this idea 
is via the definition of affine and finite morphisms. Notice that the definition of finite is a bit clunky because it requires us 
to restate the definition of affine withing the definiton of a finite morphism. Gross. We should be able to write it as \textit{``affine + $\square$.''}

For some examples, let $Y=\Spec k$. Then $X\to \Spec k$ is affine if and only if $X$ is affine. So when is $X\to\Spec k$ finite? Well since $X$ is affine, $X=\Spec A$ 
where each $A$ is a finite dimensional (reduced!) $k$-algebra. Therefore it is Artinian and we get a decomposition 
\[A=A_1\times\cdots\times A_n\]
where each $A_i$ are finite dimensional local rings with maximals $\frakm_i\subset A_i$ that are nilpotent (use Nakayama!). So by this theorem (since $A$ is reduced),
we get $A_i=k$, so 
\[A\cong k\times\cdots\times k.\]
So then $\Spec A=\sqcup \Spec k$, just a handful of points!

Let's look at another example we saw on the first day: $k[x]\to k[t]$ sending $x\mapsto t^2$. This corresponds to a map $\A1\to\A1$, 
and we can talk about the fibers of this morphism. They are all (well, almost all) of size two. But they are definitely finite! Similarly if 
we have a finite map $A\to B$, we can ask if it has finite fibers. Here if $\varphi:A\to B$ is our morphism and $f:\Spec B\to\Spec A$ is the map of rings,
then since
\[f^{-1}(\frakm)=\{\frakn\subseteq B|\varphi(n)=m\}\]
we get a lemma 
\begin{lem}
	Given a map $f:\Spec B\to \Spec A$ and $\frakm\in\Spec A$, 
	\[f^{-1}(\frakm)=\{\frakn\subseteq B/\frakm B,\text{ maximal}\}\]
\end{lem}
And so (think this through) the answer is \textbf{yes, the fibers are finite.}

A natural question involves the following 
\begin{defn}
	A morphism $f:X\to Y$ is \textbf{quasi-finite} if for all $y\in U$, $f^{-1}(y)$ is finite.
\end{defn}
and we ask: is this equivalent to finiteness? The answer is no. Consider the map $\A1\setminus 0\hookrightarrow\A1$. Both of these are affine, so the map is affine. 
But definitely the $k[t]\to k[t,t^{-1}]$ is not finite!

\begin{prop}
	Let $X\to Y$ be finite and $Z\subseteq X$ clsoed. Then $f(Z)\subseteq Y$ closed.
\end{prop}
Try it!

### 2019-10-26 Algebraic Groups/Algebraic-Groups.tex
update

>>>>

\section{October 25th, 2019}
We are going to deviate slightly from the result promised last time, because we want to talk first a bit about 
\subsection{Descent}
Recall that we had two results before about descent involving faithfully flat morphisms of finite type (where the schemes are Noetherian).

\begin{ex}Let $X=\Spec B$ and let $U_i$ be a Zariski-open cover. Then the map 
\[\sqcup U_i\to X\]
is faithfully flat and of finite type.
\end{ex}
So then one idea is that if we have a nice Zariski cover of $X$, flatness, faithful flatness, and finite type are all ``local on the source''
in that you just have to check the property locally on $X$.

\begin{prop}
	Let $A\to B$ be a faithfully flat ring homomorphism. Then 
	\begin{center}
		\begin{tikzcd}
			A\ar[r] & B\ar[r,shift left=0.75ex, "p_1"]\ar[r,shift right=0.75ex,"p_2",swap] & B\otimes _AB
		\end{tikzcd}
	\end{center}
	where $p_1(b)=b\otimes 1$ and $p_2(b)=1\otimes b$ is an exact sequence. More generally, 
	\begin{center}
		\begin{tikzcd}
			M\ar[r] & M\otimes_A B\ar[r,shift left=0.75ex,"\id\otimes p_1"]\ar[r,shift right=0.75ex,swap,"\id\otimes p_2"] & M\otimes_A B\otimes_A B
		\end{tikzcd}
	\end{center}
	is exact.
\end{prop}

There is some geometry that lost me a bit. Sorry. :(
\begin{prop}
	If $A\to B$ is faithfully flat, then the map $\rmod A\to\{(N,\alpha)|N\in\lmod B, \alpha:p_1^\ast N\cong p_2^\ast N, P(\alpha)\}$ 
	sending
	\[M\mapsto(M_B=M\otimes_AB,\alpha_{can})\]
	is 
	an equivalence of categories where $P(\alpha)$ means that $\alpha$ satisfies a cocyle condition. 
\end{prop}
\begin{rmk}
	The canonical isomorphism above is 
	\[\alpha_{can}:M_B\otimes_{B,p_1}(B\otimes_AB)\to M_B\otimes_{B,p_2}(B\otimes_A B)\]
\end{rmk}
